maps

### 2022-11-08 code/maps.qmd
Finish code for map of exome sampling points

---
title: "maps"
format: html
editor: visual
---
## Load packages
```{r}
library(sf)
library(tmap)
library(stars)
library(tidyverse)
```

## Map of sample sites for exome capture analysis
### Import data
```{r}
nps_boundary <- st_read("/home/roland/pCloudDrive/MLRG/gis/Coverages/NPS_boundary/yose_kica_sequ_boundaries.shp", stringsAsFactors=F) #read in shapefile like a dataframe using sf package
class(nps_boundary)
str(nps_boundary)
st_crs(nps_boundary) #determine CRS
utm11nCRS <- st_crs(nps_boundary)

# California Color Hillshade (90m) from koordinates.com
hillshade <- read_stars("/home/roland/Downloads/hillshade-90m-GTiff-UTMzone11/california-color-hillshade-90m.tif")
hillshade

```

### Check if shapefile polygons are valid
```{r}
st_is_valid(nps_boundary) # all 3 are invalid

nps_boundary <- st_make_valid(nps_boundary$geometry)
# tmap_options(check.and.fix = TRUE) # Don't know how to use this option, or whether it would fix the issue

st_is_valid(nps_boundary) # fixed
```

### Create map
```{r}
tm_shape(hillshade) + 
  tm_raster(style = "cont") + 
  tm_shape(nps_boundary) +
  tm_polygons()

tm_shape(nps_boundary) +
  tm_polygons()

```

>>>>

---
title: "maps"
format: html
editor: visual
---

## Load packages

```{r}
library(sf)
library(tmap)
library(stars)
library(tidyverse)
library(RPostgreSQL)
```

## Map of sample sites for exome capture analysis

Run the following 5 chunks only if you have access to the project database. For those without the necessary credentials, use the saved output file in data/maps/ that is available via the "import-map-data" chunk.

### Create dataset of sample collection points
```{r}
exome_sites <- read_csv(here::here("data", "maps", "exome_sample_sites.csv"))
```
#### Connect to database

```{r}
source("db_connect.R")
```

#### Retrieve data

```{sql retrieve-site-xy, connection=con, output.var="exome_sites_db"}
select
  id as site_id,
  utme,
  utmn
from site
where id in (74281, 50731, 50307, 52127, 10220, 10490, 50837, 20169, 72996, 72808, 70567, 54188, 10055, 10593, 50783)
```

#### Disconnect from database

```{r}
source("db_disconnect.R")
```

#### Create points layer

```{r}
exome_sites_xy <- 
  exome_sites %>% 
  inner_join(exome_sites_db, by = "site_id")

# Add x-y random noise to obscure exact site locations
exome_sites_xy <- exome_sites_xy %>% 
  add_column(x = sample(-2000:2000, size = 15),
             y = sample(-2000:2000, size = 15))

exome_sites_xy <- exome_sites_xy %>% 
  mutate(utme_noise = utme + x,
         utmn_noise = utmn + y) %>% 
  select(-utme, -utmn, -x, -y)

# Save file 
write_csv(exome_sites_xy, here::here("data", "maps", "exome_sites_xy.csv"))
```

### Import map layers

```{r import-map-data}
# Read file, if lacking database access
exome_sites_xy <- read_csv(here::here("data", "maps", "exome_sites_xy.csv"))
exome_points = st_as_sf(exome_sites_xy, coords = c("utme_noise", "utmn_noise"), crs = st_crs(26911))

nps_boundaries <- st_read(here::here(
  "data", "maps", "nps-boundaries", "yose_kica_sequ_boundaries.shp"))
st_crs(nps_boundaries) #determine coordinate reference system (CRS)

# California Color Hillshade (90m) from koordinates.com
hillshade <- read_stars(here::here(
  "data", "maps", "hillshade-90m", "california-color-hillshade-90m.tif"))
st_crs(hillshade)
```

### Check & fix shapefile polygons

```{r}
st_is_valid(nps_boundaries) # all 3 are invalid

nps_boundaries <- st_make_valid(nps_boundaries$geometry)

st_is_valid(nps_boundaries) # fix polygons
```

### Transform CRS of nps_boundary

```{r}
nps_boundaries <- st_transform(nps_boundaries, 26911)
```

### Crop raster with bounding box

```{r}
bbox <- st_bbox(c(xmin = 238000, ymax = 4235000, xmax = 403000, ymin = 4011000), crs = st_crs(nps_boundaries))
hillshade_crop <- st_crop(hillshade, bbox)
hillshade_crop <- st_transform(hillshade_crop, 26911)
```

### Create map

```{r}
exome_points_map <- tm_shape(hillshade_crop) + 
  tm_rgb(r = 1, g = 2, b = 3) +
  tm_shape(nps_boundaries) + 
  tm_borders(col = "black") +
  tm_shape(exome_points) + 
  tm_symbols(size = 0.5, 
             col = "bd_status", 
             palette = c("#FFEC8B", "#87CEFA"), # tmap only recognizes basic color names (e.g., "blue" but not "lightblue") - use HEX instead
             title.col = "Bd status") +
  tm_layout(legend.position = c("right", "top")) +
  tm_compass(type = "4star",
             position = c("left", "bottom"),
             size = 2) +
  tm_scale_bar(breaks = c(0, 25, 50), 
               text.size = 0.7, 
               position = c("left", "bottom"))

exome_points_map
```

### Save map
```{r}
tmap_save(exome_points_map, here::here("out", "map_exome_points.png"))
```
